BareTag Tool-Tracker\\ {\footnotesize {Team 3 - University of Massachusetts Amherst Senior Design Project 2025 }}

### 2024-12-19 PDR/PDR_draft_of_SDP_report/draft.tex
mdr report update

\subsection{Custom Hardware}
We designed two custom PCBs to act as the Tag and Anchors of our system. 
In our current system, we were able to trilaterally locate the custom PCB 
Tag with 3 Anchors: two Qorvo DWM3001CDK development boards and one of our 
custom PCB Anchors. Our main goal when creating our Tag PCB was to make it 
as small as possible. We eventually want these Tags to be placeable inside 
of tools. In our first revision, we were able to minimize the size of the Tag 
to approximately 26.8 by 31.6 mm squared. We were also able to successfully 
program our DWM3001C with the Edge-Connect seen in Fig 5 attached to the 
castellated edges of our Tag PCB. Because of this, we will be able to remove 
the 10-pin programming header seein in Fig. 3 to reduce the size of the Tag 
PCB in our second revision.

@@

While we were able to integrate our PCBs to make a functioning system, we 
still had some issues that we plan to address in our next revisions. The 
battery connectors into the PCBs on the Tags and the Anchors are facing 
into other components on the board, making it difficult to connect and 
disconnect our batteries to them. Our LEDs were meant to flash during 
transmission and reception of Ultra-Wideband pings on the Anchor and the 
Tag. While this was not achieved, the main functionality of our PCBs were 
not impacted. The UWB antenna on the Anchor PCB is on the opposite side of 
the LoRa antenna because we were originally worried about signals interfering 
with each other from the two different modules. In our next revision, we will 
have both antennas facing up to have them in their ideal orientation for 
transmitting and receiving signals. With this modification, we will be able to 
maximize the area where we can precisely locate our Tag with Ultra-Wideband 
pings. As you can see in Fig. 3 and Fig. 4, there are many more components on 
each of our boards. These include resistors to limit current flow into certain 
components, capacitors to reduce noise from input voltage, and LEDs that were 
meant to be used to help us debug our system. Most importantly, we had a 
voltage regulator to limit the input voltage into our Qorvo module to its 
preferred 3.3 volts. We also have many test points that helped us debug our 
system which we will remove in future revisions. Overall, our first revisions 
of our custom PCBs worked well enough for our prototype to function properly.

@@

>>>>

\subsection{Custom Hardware}
We designed two custom PCBs to act as the Tag and Anchors of our system. 
In our current system, we were able to trilaterally locate the custom PCB 
Tag with 3 Anchors: two Qorvo DWM3001CDK development boards and one of our 
custom PCB Anchors. Our main goal when creating our Tag PCB was to make it 
as small as possible. We eventually want these Tags to be placeable inside 
of tools. In our first revision, we were able to minimize the size of the Tag 
to approximately 26.8 by 31.6 mm squared. We were also able to successfully 
program our DWM3001C with the Edge-Connect seen in Fig. 4 attached to the 
castellated edges of our Tag PCB. Because of this, we will be able to remove 
the 10-pin programming header seein in Fig. 2 to reduce the size of the Tag 
PCB in our second revision.

@@

While we were able to integrate our PCBs to make a functioning system, we 
still had some issues that we plan to address in our next revisions. The 
battery connectors into the PCBs on the Tags and the Anchors are facing 
into other components on the board, making it difficult to connect and 
disconnect our batteries to them. Our LEDs were meant to flash during 
transmission and reception of Ultra-Wideband pings on the Anchor and the 
Tag. While this was not achieved, the main functionality of our PCBs were 
not impacted. The UWB antenna on the Anchor PCB is on the opposite side of 
the LoRa antenna because we were originally worried about signals interfering 
with each other from the two different modules. In our next revision, we will 
have both antennas facing up to have them in their ideal orientation for 
transmitting and receiving signals. With this modification, we will be able to 
maximize the area where we can precisely locate our Tag with Ultra-Wideband 
pings. As you can see in Fig. 2 and Fig. 3, there are many more components on 
each of our boards. These include resistors to limit current flow into certain 
components, capacitors to reduce noise from input voltage, and LEDs that were 
meant to be used to help us debug our system. Most importantly, we had a 
voltage regulator to limit the input voltage into our Qorvo module to its 
preferred 3.3 volts. We also have many test points that helped us debug our 
system which we will remove in future revisions. Overall, our first revisions 
of our custom PCBs worked well enough for our prototype to function properly.

\begin{figure}
\begin{center}
\includegraphics[scale=3]{mdr_tag_picture.png}
\caption{Populated Tag PCB}
\end{center}
\end{figure}

\begin{figure}
\begin{center}
\includegraphics[scale=3]{mdr_anchor_picture.png}
\caption{Populated Anchor PCB}
\end{center}
\end{figure}

\subsection{Prototype Functionality}
In accordance with our hardware block diagrams in Fig 2, the PCBs all work 
as intended. The modules and batteries communicate as per design. Fig. 4 
depicts the Edge-Connect programmer pins pressed to the edge of the Tag PCB. 
This method of programming proved to work, and we were successfully able to 
flash firmware to the nRF52833.

\begin{figure}
\begin{center}
\includegraphics[scale=1]{mdr_edge_connect.png}
\caption{Edge-Connect Programmer on Tag castellated edges}
\end{center}
\end{figure}

\subsection{Prototype Performance}
Our Prototype met all of our proposed MDR deliverables. We were able to 
integrate our two custom PCBs into our proposed system to track one custom 
Tag. Although the LEDs on our PCBs did not function properly, they did not 
cause any performance issues. Looking back at our complete design 
specifications, our current Prototype is well on its way to achieve them. 
For starters, we were able to track a Tag with 50 cm accuracy within a 
predefined 10 x 10 m zone. Although this is much smaller than our system’s 
final goal, we believe that we could have had a larger range if we had the 
space to do so. Our system was also capable of tracking a Tag outside the 
UWB zone using the Apple Find My network. A major performance metric that 
we evaluated is the battery life of our system. As previously stated, our 
design goal is to have an available system for ~1 year without the need for 
replacing any batteries. To test our current system’s battery life, we 
analyzed the current draw of our devices when the system was operating. 
Fig. 5 shows the power consumption of our singular Tag PCB while it is 
only transmitting Bluetooth Low Energy signals. For the majority of the 
time, the Tag is idle, but it wakes up and sends a bluetooth signal 
approximately every 5 seconds. In BLE mode, our Tag consumed 60.73 
microamps, giving it an approximate battery life of 200 days with a 
290 mAh battery attached to it. This was the expected power consumption 
for our Tag in BLE mode. While our Prototype’s battery life performed well 
in BLE mode, it did not perform close to how we want it to when sending 
Ultra-Wideband pings. In Fig. 6, the power consumption of the Tag in UWB 
mode can be seen changing as it switches from waiting to receive UWB pings 
to sending them. Our Tag in UWB mode has an average current draw of 51.63 
mA. With the 290 mAh battery our current system has attached to a Tag, its 
battery life is under 6 hours. This battery life is well below our system’s 
desired battery life, but there are many software power optimizations possible
for our system.

\begin{figure}
\begin{center}
\includegraphics[scale=1]{mdr_ble_power_consumption.png}
\caption{Power consumption of Tag in BLE mode}
\end{center}
\end{figure}

Lastly, we analyzed the power consumption of one of our Anchors. 
Similarly to the Tag, our custom PCB Anchor had an average current 
consumption of 49.23 mA. The Anchor acts as the main transmitter of the 
signals while the Tag is the main receiver. Although our Anchor has a 
similar current draw, it has a much longer battery life because our 
current Anchor has a 2200 mAh battery attached to it. We are less worried 
with the current draw of our Anchors because they are not size constrained, 
allowing us to connect many batteries to them.

\begin{figure}
\begin{center}
\includegraphics[scale=0.45]{mdr_uwb_tag_power_consumption.png}
\caption{Power consumption of Tag in UWB Mode}
\end{center}
\end{figure}

\begin{figure}
\begin{center}
\includegraphics[scale=0.45]{mdr_uwb_anchor_power_consumption.png}
\caption{Power consumption of Anchor in UWB Mode}
\end{center}
\end{figure}

\section{CONCLUSION}
As of now, the project operates in the following states:
\begin{itemize}
    \item Full programming control over Tag and Anchor rev.1 PCBs
    \item Tag lasts 4775.23 hours in BLE mode
    \item Tag lasts 5.62 hours in UWB mode
    \item Anchor lasts 44.7 hours
    \item Anchors send Tag location over LoRa to host device
    \item Find My server runs on Raspberry Pi
\end{itemize}

\section*{ACKNOWLEDGMENTS}
We thank our advisor Professor Jeremy Gummesson for the endless 
advice and support throughout the design process. We also thank 
our evaluators Professor Tilman Wolf and Professor David McLaughlin 
for essential feedback from each quarterly presentation. Of course, 
thank you to all of the course coordinators for ensuring that the 
necessary support has been available. Additionally, thank you to 
Baird Soules and M5 staff for supplying a productive engineering 
environment for prototyping and rework, and for many important 
components that we used in our project.

@@

\subsection{Technical Standards}
Bluetooth Low Energy, specified by IEEE 802.15.1, is used for our Tags to 
communicate with nearby Apple Find My ‘finder’ devices to send data to the 
Apple Find My network for global location tracking.

Ultra-Wideband is used by our Tags and Anchors for local location tracking 
with 50cm accuracy. Our application of UWB is specified in IEEE P802.15.14, 
which describes UWB used for Ad Hoc precision tracking. This spec builds on 
other specs of importance such as IEEE 802.15.4.

LoRa, discussed in IEEE spec P1451.5.5 , is used for sending data to the 
Base Device. We have chosen to not implement LoRaWAN due to the features of 
our transceiver modules, the minimal amount data we are sending, and a desire 
to not use additional hardware such as a LoRaWAN gateway.

Our chosen RF emitters all adhere to FCC guidelines as defined in FCC 47 
CFR Part 15, which states that any radiator must be licensed unless it
meets 47 CFR 15, unless otherwise given specific approval by the FCC. 
FCC certifications can be found on the product pages for all of our 
hardware radiators.

\subsection{Testing Methods}
Apple Find My integration has been tested through validating location 
data displayed on a locally hosted GUI provided by [9].

UWB tracking has been tested with 3 Anchors and 1 Tag as of MDR. 50cm 
accuracy has been noted. This was performed through the manual creation 
of a coordinate system in an indoor open-field environment, and moving 
the Tag to different positions in the grid.

\subsection{Project Expenditures}
\begin{center}
\includegraphics[scale=0.35]{mdr_project_expenditures.png}
\end{center}

\subsection{Project Management}
Team communication is done through text and Discord. We meet on a weekly 
basis both with each other and Professor Gummeson. GitHub and Google Drive 
are used as our version control platforms. Each team member has a unique 
skill-set that makes up for what another lacks, making the team dynamic 
optimal for a complex project. The experience of our team is evenly balanced 
with backgrounds in hardware and software, which is reflected in the 
designated project roles.

\begin{itemize}
    \item Sean Brown - Logistics Lead and Hardware Support
    \item Connor McGarry - Hardware Lead
    \item Ken Su - Budget Lead and Software Support
    \item Walter Tebbetts - Software Lead
\end{itemize}

Varying levels of class workload makes it difficult for progress reports 
occasionally, but we try to log our progress for team review as best as 
possible. Each team member exercises leadership skills when another is too 
busy with external responsibilities to work on their designated part of the 
project for a while. Each team member understanding every aspect of the 
project has been important in helping each other out and making up for 
occasional slow progress. The largest hurdle we have overcome thus far is 
that of individual motivation and understanding of a team project. Some of 
us had not worked on a project of this scale before and therefore have 
needed to learn how to make careful decisions that affect other team members. 
Taking accountability in a team is another skill we are learning, as there 
have been a number of errors made where the best solution was to admit to 
the issue as soon as possible, and then collaborate on how to solve it.

\subsection{Beyond the Classroom}
Classes that have been essential to this project so far have been ECE523, 
ECE325, ECE304 and ECE545. These classes relate to embedded system power 
management and networking, both of which are integral to the design of our 
project. We plan on having discussions with the professors of these classes 
as we progress in our development.

We have had to do more detailed data sheet analysis than what we did in 
ECE304, and Digikey has been an essential resource for component sourcing 
and information retrieval.

We have used a variety of unfamiliar software and hardware in the project, 
all of which have required detailed research and experimentation. Since we 
have taken inspiration from GtHub projects, we have also spent a lot of time 
analyzing unfamiliar code to make sure we understand what we are using, and 
ensuring we are able to edit it according to our desired optimizations.
